# 000
---
title: "00.data"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

# library
```{r}
Sys.setenv(LANGUAGE = "en")
options(warn = -1)
memory.limit(size = 64670)
library(Seurat)
library(harmony)
library(MAST)
library(dplyr)
library(tidyselect)
library(RColorBrewer)
library(future)
library(ggplot2)
library(org.Mm.eg.db)
library(EnsDb.Mmusculus.v79)
library(cowplot)
library(data.table)
library(clusterProfiler)
#library(DoubletDecon)
library(DoubletFinder)
```

# paralelization set 
```{r}
# plan("multiprocess", workers = 4)
# options(future.globals.maxSize = 1024^4)
# plan()
# plan("sequential")
```

# theme set
```{r}
theme_set(theme_cowplot(font_size = 8))
theme.text = theme_cowplot(font_size = 8)
cols = c(brewer.pal(9, "Set1"),
         brewer.pal(8, "Set2")[-c(2,4,8)],
         brewer.pal(12, "Set3")[-9],
         brewer.pal(12, "Paired"),
         brewer.pal(8, "Dark2"),
         brewer.pal(11, "Spectral")[-6],
         brewer.pal(11, "BrBG")[-6]
         )
```

# load data
```{r}
load("01.sct.dnt.Rdata")
load("01.sct.nk.Rdata")
load("01.sct.sp.Rdata")

load("01.sct.kept.Rdata")

load("01.harmony.dnt.nk.Rdata")
load("01.harmony.dnt.nk.meta.Rdata")

# load("01.harmony.dnt.Rdata")
# load("01.harmony.dnt.cc.marker.Rdata")

load("01.FindMarkers.between.DNT.and.CD4.CD8.NK.Rdata")
load("01.FindMarkers.between.DNT.and.CD4.CD8.NK.join.Rdata")

load("01.harmony.all.Rdata")
load("01.harmony.all.meta.all.Rdata")

load("01.cca.kept.Rdata")
load("01.cca.kept.cc.marker.Rdata")
load("01.cca.kept.cc.marker.48nk.Rdata")

load("01.cca.unfilter.Rdata")
load("01.cca.unfilter.meta.all.Rdata")

# load("01.cc.marker.DNT.Rdata")
# load("01.cca.kept.cc.marker.DNT.48nk.Rdata")

load("01.nDNT.decon.Rdata")
load("01.DoubletDecon.DoubletFinder.Rdata")

load("01.trans.marker.Rdata")
```

# prepare data (sct)
```{r}
##### run sct #####
sct = mapply(function(sample, data.dir){
    print(paste("Preparing", sample))
    ##### read the 10x files of different formats #####
    if (sample %in% c("CD4", "CD8", "TCRab", "nDNT_rep1", "nDNT_rep2", "aDNT_rep1", "aDNT_rep2")) 
      {
      sct = Read10X(data.dir = data.dir)
      } 
    else if (sample %in% c("NK_sp1", "NK_sp2", "NK_sp3")) 
      {
      sct = data.table::fread(file = list.files(data.dir, full.names = T), header = "auto", sep = ",") %>%
      tibble::column_to_rownames("V1")
      } 
    else if (sample %in% c("sp3_rep1", "sp3_rep2", "sp4_rep1", "sp4_rep2"))
      {
      sct = data.table::fread(file = data.dir, header = "auto", sep = ",") %>%
      dplyr::mutate(SYMBOL = AnnotationDbi::mapIds(x = EnsDb.Mmusculus.v79, keys = V1, column = "SYMBOL", keytype = "GENEID")) %>%
      na.omit() %>%
      dplyr::filter(!SYMBOL == "") %>%
      dplyr::select(-V1)
      dup = sct$SYMBOL[duplicated(sct$SYMBOL)]
      z1 = sct %>%
      dplyr::filter(SYMBOL %in% dup) %>%
      dplyr::group_by(SYMBOL) %>%
      dplyr::summarise_all(sum)
      # long time, but should work
      z2 = sct %>%
      dplyr::filter(!SYMBOL %in% dup)
      sct = rbind(z1, z2) %>%
      tibble::column_to_rownames("SYMBOL")
      }
    print(head(sct[,1:5]))
    print(head(sct[,(ncol(sct) - 5):ncol(sct)]))
    
    ##### create the seurat objects #####
    sct = sct %>%
    CreateSeuratObject(project = sample) %>%
    PercentageFeatureSet(pattern = "^mt-", col.name = "percent.mt") %>%
    RenameCells(add.cell.id = sample)
    print(dim(sct)) # print the cell number before qc
    
    ##### subset the seurat objects #####
    sct = subset(sct, subset = percent.mt < 10 & nFeature_RNA > 500)
    print(dim(sct)) # print the cell number after qc
    print(median(sct$nCount_RNA)) # print the median UMIs after qc
    print(max(sct$nCount_RNA)) # print the max UMIs after qc
    
    ##### scale each samples #####
    sct = SCTransform(object = sct, 
                    vars.to.regress = "percent.mt", 
                    verbose = F,
                    conserve.memory = T )
    
    return(sct)
    }, 
  sample = c(list.files(path = "../data/")), 
  data.dir = c(list.files(path = "../data/", full.names = T)),
  SIMPLIFY = F)

##### split and save the sct files #####
sct.dnt = sct[1:7]
sct.nk = sct[8:10]
sct.sp = sct[11:14]
save(sct.dnt, file = "../result/01.sct.dnt.Rdata")
save(sct.nk,  file = "../result/01.sct.nk.Rdata")
save(sct.sp,  file = "../result/01.sct.sp.Rdata")
rm(sct)
gc()

##### sct.kept #####
sct.kept = mapply(function(sct, sample){
    DefaultAssay(sct) = "RNA"
    print(paste("Preparing", sample))
    ###### subset ######
    if (sample %in% c("nDNT_rep1", "nDNT_rep2", "aDNT_rep1", "aDNT_rep2")) 
      {
      sct.kept = subset(x = sct,
                        subset = (Cd3d>0|Cd3e>0|Cd3g>0)&Cd4==0&Cd8b1==0&Klrb1c==0,
                        slot = "counts")### layer data ###
      } 
    else if (sample %in% "CD4") 
      {
      sct.kept = subset(x = sct, subset = (Cd3d>0|Cd3e>0|Cd3g>0)&Cd8a==0&Cd8b1==0&Klrb1c==0 )
      } 
    else if (sample %in% "CD8") 
      {
      sct.kept = subset(x = sct, subset = (Cd3d>0|Cd3e>0|Cd3g>0)&Cd4==0&Klrb1c==0 )
      }
    else if (sample %in% "TCRab") 
      {
      sct.kept = subset(x = sct, subset = (Cd3d>0|Cd3e>0|Cd3g>0) )
      } 
    else if (sample %in% c("NK_sp1", "NK_sp2", "NK_sp3")) 
      {
      sct.kept = subset(x = sct, subset = Cd3d == 0 & Cd3g == 0)
      }
    ######
    sct.kept = SCTransform(object = sct.kept, 
                         vars.to.regress = "percent.mt", 
                         verbose = F,
                         conserve.memory = F )
    ######
    return(sct.kept)}, 
  sct = c(sct.dnt, sct.nk), 
  sample = c(names(sct.dnt), names(sct.nk)),
  SIMPLIFY = F)

save(sct.kept, file = "../result/01.sct.kept.Rdata")
```

# harmony CD4 CD8 NK DNT kept cells
```{r run CD4 CD8 NK DNT harmony}
##### merge the sct files #####
harmony.dnt.nk = merge(x = sct.kept$aDNT_rep1, y = sct.kept[2:10])
rm(sct.kept)
gc()
plan("sequential")
harmony.dnt.nk = harmony.dnt.nk %>%
  SCTransform(vars.to.regress = "percent.mt", conserve.memory = F) %>% 
  # take a lot memory, about 40 min.
  RunPCA(verbose = F)
##### change the core #####
plan("multiprocess", workers = 4)
options(future.globals.maxSize = 1024^4)
##### set the batch #####
table(harmony.dnt.nk$orig.ident)
harmony.dnt.nk$batch = factor(
  harmony.dnt.nk$orig.ident,
  labels = c("b1", "b2", rep("b1", 3), "b2", rep("b3", 3), "b1"))
table(harmony.dnt.nk$batch, harmony.dnt.nk$orig.ident)
##### run the harmony #####
harmony.dnt.nk = harmony.dnt.nk %>%
  RunHarmony(group.by.vars = "batch", assay.use = "SCT") %>%
  RunUMAP(dims = 1:30, reduction = "harmony") %>%
  FindNeighbors(dims = 1:30, reduction = "harmony") %>%
  FindClusters(resolution = .1)
##### set idents to be celltype #####
harmony.dnt.nk$celltype = sub("(_rep[1-2])|(_sp[1-3])", "", harmony.dnt.nk$orig.ident)
harmony.dnt.nk$celltypesplit = sub("_sp[1-3]", "", harmony.dnt.nk$orig.ident)
table(harmony.dnt.nk$celltype)
table(harmony.dnt.nk$celltypesplit)
Idents(harmony.dnt.nk) = "celltype"
##### ScaleData for findmarkers #####
plan("sequential")
DefaultAssay(harmony.dnt.nk) = "RNA"
harmony.dnt.nk = harmony.dnt.nk %>%
  NormalizeData(., normalization.method = "LogNormalize", scale.factor = 10000) %>%
  FindVariableFeatures(., selection.method = "vst", nfeatures = 2000) %>%
  ScaleData(., vars.to.regress = "percent.mt")
##### save the results #####
save(harmony.dnt.nk, file = "01.harmony.dnt.nk.Rdata")
```
## markers
```{r}
object = harmony.dnt.nk
##### FindMarkers between DNT and CD4, CD8, NK respectively #####
markers = mapply(function(ident.1, ident.2){
  FindMarkers(object = object, ident.1 = ident.1, ident.2 = ident.2, group.by = "celltypesplit", assay = "RNA", slot = "data", only.pos = T) %>%
    tibble::rownames_to_column("gene") %>%
    dplyr::mutate(type = paste(ident.1, "vs", ident.2, sep = ".")) %>%
    dplyr::arrange(desc(avg_logFC))
},
ident.1 = rep(c("nDNT_rep1", "nDNT_rep2", "aDNT_rep1", "aDNT_rep2"), 3), 
ident.2 = rep(c("CD4", "CD8", "NK"), c(4, 4, 4)),
SIMPLIFY = F, USE.NAMES = F)
names(markers) = paste(rep(c("nDNT_rep1", "nDNT_rep2", "aDNT_rep1", "aDNT_rep2"), 3), "vs", rep(c("CD4", "CD8","NK"), c(4,4,4)), sep = ".")
##### save results #####
save(markers, file = "01.FindMarkers.between.DNT.and.CD4.CD8.NK.Rdata")
##### join the replication result #####
genelist = list(
  nDNT.vs.CD4 = intersect(markers$nDNT_rep1.vs.CD4$gene, markers$nDNT_rep2.vs.CD4$gene),
  nDNT.vs.CD8 = intersect(markers$nDNT_rep1.vs.CD8$gene, markers$nDNT_rep2.vs.CD8$gene),
  nDNT.vs.NK  = intersect(markers$nDNT_rep1.vs.NK$gene,  markers$nDNT_rep2.vs.NK$gene ),
  aDNT.vs.CD4 = intersect(markers$aDNT_rep1.vs.CD4$gene, markers$aDNT_rep2.vs.CD4$gene),
  aDNT.vs.CD8 = intersect(markers$aDNT_rep1.vs.CD8$gene, markers$aDNT_rep2.vs.CD8$gene),
  aDNT.vs.NK  = intersect(markers$aDNT_rep1.vs.NK$gene,  markers$aDNT_rep2.vs.NK$gene )
                )
markers.both = mapply(function(ident.1, ident.2){
  FindMarkers(object = object, ident.1 = ident.1, ident.2 = ident.2, group.by = "celltype", assay = "RNA", slot = "data", only.pos = T) %>%
    tibble::rownames_to_column("gene") %>%
    dplyr::mutate(type = paste(ident.1, "vs", ident.2, sep = ".")) %>%
    dplyr::arrange(desc(avg_logFC))
}, 
ident.1 = rep(c("nDNT", "aDNT"),    3), 
ident.2 = rep(c("CD4", "CD8","NK"), c(2,2,2)),
SIMPLIFY = F, USE.NAMES = F)
names(markers.both) = paste(rep(c("nDNT", "aDNT"), 3), "vs", rep(c("CD4", "CD8","NK"),c(2,2,2)), sep = ".")
markers.join = lapply(names(markers.both), function(i){
  markers.both[[i]] %>% 
    dplyr::filter(gene %in% genelist[[i]])
})
names(markers.join) = names(markers.both)
##### save results #####
save(genelist, markers.both, markers.join, 
     file = "01.FindMarkers.between.DNT.and.CD4.CD8.NK.join.Rdata")
##### write the csv #####
rbind(markers.join$nDNT.vs.CD4, markers.join$nDNT.vs.CD8, markers.join$nDNT.vs.NK) %>%
  write.csv(file = "TableS2.nDNT.vs.CD4.CD8.NK.csv", x = ., row.names = F)
rbind(markers.join$aDNT.vs.CD4, markers.join$aDNT.vs.CD8, markers.join$aDNT.vs.NK) %>%
  write.csv(file = "TableS3.aDNT.vs.CD4.CD8.NK.csv", x = ., row.names = F)
##### 
```
# harmony DNT kept cells
```{r run nDNT, aDNT harmony [NOT USED]}
names(sct.kept)
plan("sequential")
harmony.dnt = mapply(function(i, j, lambda, theta, resolution){
  ##### run the harmony #####
  object = merge(i, j) %>%
    SCTransform(vars.to.regress = "percent.mt", conserve.memory = F) %>% 
    RunPCA(verbose = F) %>%
    RunHarmony(group.by.vars = "orig.ident", assay.use = "SCT",
               lambda = lambda, theta = theta) %>%
    RunUMAP(dims = 1:30, reduction = "harmony") %>%
    FindNeighbors(dims = 1:30, reduction = "harmony") %>%
    FindClusters(resolution = resolution)
  ##### ScaleData for findmarkers #####
  DefaultAssay(object) = "RNA"
  object = object %>%
    NormalizeData(., normalization.method = "LogNormalize", scale.factor = 10000) %>%
    FindVariableFeatures(., selection.method = "vst", nfeatures = 2000) %>%
    ScaleData(., vars.to.regress = "percent.mt")
  ##### return object #####
  return(object)
}, SIMPLIFY = F, 
i = list("nDNT" = sct.kept$nDNT_rep1, "aDNT" = sct.kept$aDNT_rep1), 
j = list("nDNT" = sct.kept$nDNT_rep2, "aDNT" = sct.kept$aDNT_rep2), 
lambda = c(1, .05), 
theta = c(2, 3), 
resolution = c(.05, .01))
##### run the harmony of naDNT #####
# harmony.naDNT = merge(sct.kept$aDNT_rep1, sct.kept[c(2,5:6)]) %>%
#     SCTransform(vars.to.regress = "percent.mt", conserve.memory = F) %>% 
#     RunPCA(verbose = F)
# harmony.naDNT$batch = sub("[n|a]DNT_", "", harmony.naDNT$orig.ident)
# table(harmony.naDNT$batch, harmony.naDNT$orig.ident)
# harmony.naDNT = harmony.naDNT %>%
#     RunHarmony(group.by.vars = "batch", assay.use = "SCT") %>%
#     RunUMAP(dims = 1:30, reduction = "harmony") %>%
#     FindNeighbors(dims = 1:30, reduction = "harmony") %>%
#     FindClusters(resolution = .1)
# DimPlot(harmony.naDNT, split.by = "orig.ident", ncol = 2, label = T, cols = "Paired")
# FeaturePlot(harmony.naDNT, c("Il17a","Cxcr6", "Eomes", "Gzmb", "Cd74"), order = T,label = T)
##### save the results #####
save(harmony.dnt, file = "01.harmony.dnt.Rdata")
```
## markers
```{r [NOT USED]}
cc.marker = mapply(function(object){
  lapply(sort(unique(object$seurat_clusters)), function(i){
    if (min(table(object$orig.ident[object$seurat_clusters == i])) > 30) {
      FindConservedMarkers(object, ident.1 = i, grouping.var = "orig.ident", 
                           assay = "RNA", slot = "data", only.pos = T) %>%
        tibble::rownames_to_column("gene") %>%
        dplyr::filter(max_pval < .05 & minimump_p_val < .05) %>%
        dplyr::mutate(cluster = i)
    }}) %>% 
    data.table::rbindlist() %>%
    dplyr::left_join(y = FindAllMarkers(object, assay = "RNA", slot = "data", only.pos = T), by = c("cluster", "gene")) %>%
    dplyr::arrange(desc(avg_logFC))
}, object = harmony.dnt, SIMPLIFY = F)
names(cc.marker)
save(cc.marker, file = "01.harmony.dnt.cc.marker.Rdata")
```
### 48nk.marker
```{r [NOT USED]}
object = harmony.dnt.nk
head(harmony.dnt$nDNT@meta.data)
df1 = rbind(harmony.dnt$nDNT@meta.data %>% tibble::rownames_to_column("cells") %>% dplyr::select(cells, seurat_clusters_before = seurat_clusters),
            harmony.dnt$aDNT@meta.data %>% tibble::rownames_to_column("cells") %>% dplyr::select(cells, seurat_clusters_before = seurat_clusters))
table(df1$seurat_clusters)
head(df1)
df2 = cca.kept$DNT@meta.data %>% tibble::rownames_to_column("cells") %>% dplyr::select(cells, integrated_snn_res.0.01)
table(df2$integrated_snn_res.0.01)
meta = object@meta.data %>%
  tibble::rownames_to_column("cells") %>%
  dplyr::left_join(y = df1, by = "cells") %>%
  dplyr::left_join(y = df2, by = "cells") %>%
  tibble::column_to_rownames("cells")
head(meta)
meta$cluster = paste0(meta$celltype, meta$seurat_clusters_before) %>% sub("NA", "", .) %>%
  factor(., levels = c(paste0("nDNT", 0:4), paste0("aDNT", 0:1), "CD4", "CD8", "NK", "TCRab"))
table(meta$cluster)
meta$recluster = paste0(sub("^[n|a]", "", meta$celltype), meta$integrated_snn_res.0.01) %>% sub("NA", "", .) %>%
  factor(., levels = c(paste0("DNT", 0:1), "CD4", "CD8", "NK", "TCRab"))
table(meta$recluster)
save(meta, file = "01.harmony.dnt.nk.meta1.Rdata")

# object@meta.data = meta
# cc.marker.48nk = mapply(function(ident.1, group.by, cc.marker){
#   
#   x = mapply(function(ident.2){
#     FindMarkers(object, ident.1, ident.2, group.by = group.by, only.pos = T, assay = "RNA", slot = "data",
#                 features = cc.marker) %>%
#       tibble::rownames_to_column("gene") %>%
#       dplyr::filter(p_val_adj < .05)
#   }, SIMPLIFY = F,
#   ident.2 = c("CD4", "CD8", "NK"))
#   dplyr::inner_join(x$CD4, x$CD8, by = "gene", suffix = c(".CD4", ".CD8")) %>%
#     dplyr::inner_join(., x$NK, by = "gene", suffix = c("", ".NK")) %>%
#     dplyr::mutate(type = ident.1)
#   
# }, SIMPLIFY = F, 
# ident.1 = c(paste0("nDNT",0:4), paste0("aDNT",0:1), paste0("DNT",0:1)),
# group.by = c(rep("cluster",7), rep("recluster",2)),
# cc.marker = sapply(cc.marker, function(i){
#   sapply(sort(unique(i$cluster)), function(j){
#     (i %>% dplyr::filter(cluster == j))$gene
#     })
#   }) %>% 
#   unlist(., recursive = F)) %>%
#   data.table::rbindlist()
# head(cc.marker.48nk)
# save(cc.marker.48nk, file = "01.cca.kept.cc.marker.48nk.Rdata")
```
# harmony all
```{r run harmony all}
##### merge sct files #####
harmony.all = merge(x = sct.dnt$aDNT_rep1, y = c(sct.dnt[2:7], sct.nk, sct.sp))
rm(sct.dnt, sct.nk, sct.sp)
gc()
plan("sequential")
harmony.all = harmony.all %>%
  SCTransform(vars.to.regress = "percent.mt", conserve.memory = F) %>% 
  # take a lot memory, about 40 min.
  RunPCA(verbose = F)
##### change the core #####
plan("multiprocess", workers = 4)
options(future.globals.maxSize = 1024^4)
##### set the batch #####
table(harmony.all$orig.ident)
harmony.all$batch = factor(
  harmony.all$orig.ident,
  labels = c("b1", "b2", rep("b1", 3), "b2", rep("b3", 3), rep("b4", 4), "b1"))
table(harmony.all$batch, harmony.all$orig.ident)
##### run the harmony #####
harmony.all@meta.data = meta.all
harmony.all = RunHarmony(object = harmony.all, group.by.vars = "batch", assay.use = "SCT", lambda = .1, theta = 3) %>%
  RunUMAP(dims = 1:30, reduction = "harmony") %>%
  FindNeighbors(dims = 1:30, reduction = "harmony") %>%
  FindClusters(resolution = .1)
DimPlot(harmony.all, group.by = "celltype.converge", label = T, cols = cols)
##### set idents to be celltype #####
harmony.all$celltype = factor(
  harmony.all$orig.ident,
  labels = c("aDNT", "aDNT", "CD4", "CD8", "nDNT", "nDNT",
             rep("NK", 3),
             rep("sp", 4),
             "TCRab"))
harmony.all$celltypesplit = factor(
  harmony.all$orig.ident,
  labels = c("aDNT_rep1", "aDNT_rep2", "CD4", "CD8", "nDNT_rep1", "nDNT_rep2",
             rep("NK", 3),
             rep("sp", 4),
             "TCRab"))
table(harmony.all$celltype)
table(harmony.all$celltypesplit)
Idents(harmony.all) = "celltype"
##### ScaleData for findmarkers #####
plan("sequential")
DefaultAssay(harmony.all) = "RNA"
harmony.all = harmony.all %>%
  NormalizeData(., normalization.method = "LogNormalize", scale.factor = 10000) %>%
  FindVariableFeatures(., selection.method = "vst", nfeatures = 2000) %>%
  ScaleData(., vars.to.regress = "percent.mt")
##### save the results #####
save(harmony.all, file = "01.harmony.all.Rdata")
```
```{r modify harmony all metadata}
##### save the pre metadata #####
metapre = harmony.all@meta.data
save(metapre, file = "01.harmony.all.metapre.Rdata")
##### import the markers from plos dataset #####
plos.meta1 = fread(input = "../01.processed_data/mouse_spleen_2019_plos.biology/other information/pbio.3000528.s013.csv") # Mouse spleen scRNA-seq clustering data.
##### modify the plos metadata #####
meta.plos = mapply(function(i, j){
  plos.meta1[grep(i, plos.meta1$cell), ] %>%
    dplyr::mutate(cell.rename = paste(j, cell, sep = "_"))
  },
  i = c("mouse_3.1", "mouse_3.2", "mouse_4.1", "mouse_4.2"), 
  j = c("sp3_rep1", "sp3_rep2", "sp4_rep1", "sp4_rep2"),
  SIMPLIFY = F) %>% 
  rbindlist() %>%
  dplyr::mutate(converged.cell.type.simple = sub(".[1-4]$", "", converged.cell.type)) %>%
  dplyr::select(cell.rename, converged.cell.type.simple)
head(meta.plos)
colnames(meta.plos)
table(meta.plos$converged.cell.type.simple)
##### gene expression #####
DefaultAssay(harmony.all) = "RNA"
meta.gene = FetchData(object = harmony.all, c("Cd3d", "Cd3e", "Cd3g", "Cd4", "Cd8a", "Cd8b1", "Klrb1c"), slot = "data") %>%
  tibble::rownames_to_column("cell.rename")
head(meta.gene)
##### merge the pre metadata with DNT and plos metadata #####
meta.all = metapre %>%
  tibble::rownames_to_column("cell.rename") %>%
  # dplyr::left_join(y = meta.DNT,  by = c("cell.rename")) %>%
  dplyr::left_join(y = meta.plos, by = c("cell.rename")) %>%
  dplyr::left_join(y = meta.gene, by = c("cell.rename")) %>%
  tibble::column_to_rownames("cell.rename")
head(meta.all)
##### put celltype inside converged.cell.type.simple #####
meta.all$converged.cell.type.simple[is.na(meta.all$converged.cell.type.simple)] =
  as.character(meta.all$celltypesplit)[is.na(meta.all$converged.cell.type.simple)]
table(meta.all$converged.cell.type.simple)
##### celltype.converge #####
meta.all$celltypesplit.converge = sub("_rep[1|2]", "", meta.all$converged.cell.type.simple)
meta.all$celltype.converge = sub("_rep[1|2].*", "", meta.all$converged.cell.type.simple)
table(meta.all$celltypesplit.converge)
table(meta.all$celltype.converge)
##### T identity #####
meta.all$Tid = (meta.all$Cd3d > 0 |meta.all$Cd3e > 0| meta.all$Cd3g > 0) & 
  meta.all$celltype.converge %in% c("aDNT", "CD4", "CD8", "nDNT", "TCRab",
                                    "Memory-CD4-T-cell", "Memory-CD8-T-cell",
                                    "Naive-CD4-T-cell", "Naive-CD8-T-cell",
                                    "NK-T-cell")
table(meta.all$Tid, meta.all$celltype.converge)

meta.all$CD4id = (meta.all$Cd3d > 0 |meta.all$Cd3e > 0| meta.all$Cd3g > 0) &
  meta.all$Cd8a == 0 & meta.all$Cd8b1 == 0 & meta.all$Klrb1c == 0 &
  meta.all$celltype.converge %in% c("CD4",
                                    "Memory-CD4-T-cell",
                                    "Naive-CD4-T-cell")
table(meta.all$CD4id, meta.all$celltype.converge)

meta.all$CD8id = (meta.all$Cd3d > 0 |meta.all$Cd3e > 0| meta.all$Cd3g > 0) &
  meta.all$Cd4 == 0 & meta.all$Klrb1c == 0 & 
  meta.all$celltype.converge %in% c("CD8",
                                    "Memory-CD8-T-cell",
                                    "Naive-CD8-T-cell")
table(meta.all$CD8id, meta.all$celltype.converge)

meta.all$DNTid = (meta.all$Cd3d > 0 |meta.all$Cd3e > 0| meta.all$Cd3g > 0) &
  meta.all$Cd4 == 0 & meta.all$Klrb1c == 0 & meta.all$Cd8b1 == 0 &
  meta.all$celltype.converge %in% c("aDNT", "nDNT")
table(meta.all$DNTid, meta.all$celltype.converge)

meta.all$NKid = (meta.all$Cd3d == 0 & meta.all$Cd3g == 0) &
   meta.all$celltype.converge %in% c("NK", "NK-cell")
table(meta.all$NKid, meta.all$celltype.converge)

meta.all$contaminate = (meta.all$Tid == FALSE & 
  meta.all$celltype.converge %in% c("aDNT", "CD4", "CD8", "nDNT", "TCRab",
                                    "Memory-CD4-T-cell", "Memory-CD8-T-cell",
                                    "Naive-CD4-T-cell", "Naive-CD8-T-cell",
                                    "NK-T-cell")) |
  (meta.all$CD4id == FALSE &
  meta.all$celltype.converge %in% c("CD4",
                                    "Memory-CD4-T-cell",
                                    "Naive-CD4-T-cell")) |
  (meta.all$CD8id == FALSE & 
  meta.all$celltype.converge %in% c("CD8",
                                    "Memory-CD8-T-cell",
                                    "Naive-CD8-T-cell")) |
  (meta.all$DNTid == FALSE & 
  meta.all$celltype.converge %in% c("aDNT", "nDNT")) |
  (meta.all$NKid == FALSE &
  meta.all$celltype.converge %in% c("NK", "NK-cell"))
table(meta.all$contaminate, meta.all$celltype.converge)
##### save modified metadata #####
save(meta.all, file = "01.harmony.all.meta.all.Rdata")

```
# cca kept
```{r}
plan("sequential")
cca.kept = mapply(function(object.list){
  anchor.features = SelectIntegrationFeatures(object.list = object.list, 
                                              nfeatures = 3000) 
  object.integrated = PrepSCTIntegration(object.list = object.list, 
                                         anchor.features = anchor.features) %>%
    lapply(FUN = RunPCA, features = anchor.features, verbose = F) %>%
    FindIntegrationAnchors(object.list = ., 
                           reduction = "rpca",
                           anchor.features = anchor.features, 
                           normalization.method = "SCT",
                           reference = 1) %>% # long time
    IntegrateData(anchorset = ., 
                  normalization.method = "SCT") %>%
    RunPCA(object = .) %>%
    RunUMAP(object = ., 
            dims = 1:30) %>%
    FindNeighbors(dims = 1:30) %>%
    FindClusters(resolution = .05)
  ##### scale for finding markers #####
  DefaultAssay(object.integrated) = "RNA"
  object.integrated = NormalizeData(object.integrated,
                                    normalization.method = "LogNormalize",
                                    scale.factor = 10000)
  object.integrated = FindVariableFeatures(object.integrated,
                                           selection.method = "vst",
                                           nfeatures = 2000)
  object.integrated = ScaleData(object.integrated,
                                vars.to.regress = "percent.mt")
  return(object.integrated)
}, 
object.list = list("nDNT" = sct.kept[5:6], "aDNT" = sct.kept[1:2], "DNT" = sct.kept[c(5:6, 1:2)]), SIMPLIFY = F)
DefaultAssay(cca.kept$DNT) = "integrated"
cca.kept$DNT = FindClusters(cca.kept$DNT, resolution = .01)
save(cca.kept, file = "01.cca.kept.Rdata")
```
## markers
```{r}
cc.marker = mapply(function(object){
  print(table(object$seurat_clusters, object$orig.ident))
  lapply(sort(unique(object$seurat_clusters)), function(i){
    if (min(table(object$orig.ident[object$seurat_clusters == i])) > 30){
      FindConservedMarkers(object, ident.1 = i, grouping.var = "orig.ident", 
                           assay = "RNA", slot = "data", only.pos = T) %>%
        tibble::rownames_to_column("gene") %>%
        dplyr::filter(max_pval < .05 & minimump_p_val < .05) %>%
        dplyr::mutate(cluster = i)
    }}) %>% 
    data.table::rbindlist() %>%
    dplyr::left_join(y = FindAllMarkers(object, assay = "RNA", slot = "data", only.pos = T), by = c("cluster", "gene")) %>%
    dplyr::group_by(cluster) %>%
    dplyr::arrange(desc(avg_logFC), .by_group = T)
}, object = cca.kept, SIMPLIFY = F)
names(cc.marker)
save(cc.marker, file = "01.cca.kept.cc.marker.Rdata")
```
### 48nk.marker
```{r}
object = harmony.dnt.nk
# df1 = rbind(cca.kept$nDNT@meta.data %>% tibble::rownames_to_column("cells") %>% dplyr::select(cells, integrated_snn_res.0.05), 
#             cca.kept$aDNT@meta.data %>% tibble::rownames_to_column("cells") %>% dplyr::select(cells, integrated_snn_res.0.05))
# table(df1$integrated_snn_res.0.05)
# head(df1)
# df2 = cca.kept$DNT@meta.data %>% tibble::rownames_to_column("cells") %>% dplyr::select(cells, integrated_snn_res.0.01)
# table(df2$integrated_snn_res.0.01)
# meta = object@meta.data %>%
#   tibble::rownames_to_column("cells") %>%
#   dplyr::left_join(y = df1, by = "cells") %>%
#   dplyr::left_join(y = df2, by = "cells") %>%
#   tibble::column_to_rownames("cells")
# head(meta)
# meta$cluster = paste0(meta$celltype, meta$integrated_snn_res.0.05) %>% sub("NA", "", .) %>%
#   factor(., levels = c(paste0("nDNT", 0:4), paste0("aDNT", 0:1), "CD4", "CD8", "NK", "TCRab"))
# table(meta$cluster)
# meta$recluster = paste0(sub("^[n|a]", "", meta$celltype), meta$integrated_snn_res.0.01) %>% sub("NA", "", .) %>%
#   factor(., levels = c(paste0("DNT", 0:1), "CD4", "CD8", "NK", "TCRab"))
# table(meta$recluster)
# save(meta, file = "01.harmony.dnt.nk.meta.Rdata")

object@meta.data = meta
cc.marker.48nk = mapply(function(ident.1, group.by, cc.marker){
  
  x = mapply(function(ident.2){
    FindMarkers(object, ident.1, ident.2, group.by = group.by, only.pos = T, assay = "RNA", slot = "data",
                features = cc.marker) %>%
      tibble::rownames_to_column("gene") %>%
      dplyr::filter(p_val_adj < .05)
  }, SIMPLIFY = F,
  ident.2 = c("CD4", "CD8", "NK"))
  dplyr::inner_join(x$CD4, x$CD8, by = "gene", suffix = c(".CD4", ".CD8")) %>%
    dplyr::inner_join(., x$NK, by = "gene", suffix = c("", ".NK")) %>%
    dplyr::mutate(type = ident.1)
  
}, SIMPLIFY = F, 
ident.1 = c(paste0("nDNT",0:4), paste0("aDNT",0:1), paste0("DNT",0:1)),
group.by = c(rep("cluster",7), rep("recluster",2)),
cc.marker = sapply(cc.marker, function(i){
  sapply(sort(unique(i$cluster)), function(j){
    (i %>% dplyr::filter(cluster == j))$gene
    })
  }) %>% 
  unlist(., recursive = F)) %>%
  data.table::rbindlist()
head(cc.marker.48nk)
save(cc.marker.48nk, file = "01.cca.kept.cc.marker.48nk.Rdata")
```
### up and downregulate
```{r}
object = cca.kept$DNT
object$celltype = sub("_rep[1|2]", "", object$orig.ident) %>% factor(., levels = c("nDNT", "aDNT"))
table(object$integrated_snn_res.0.01, object$orig.ident)
DefaultAssay(object) = "RNA"
df = mapply(function(ident.1, ident.2, subset.ident){
  x = FindMarkers(object, ident.1 = ident.1, ident.2 = ident.2, subset.ident = subset.ident, 
                  group.by = "orig.ident", assay = "RNA", slot = "data") %>%
    tibble::rownames_to_column("gene") %>%
    dplyr::filter(p_val_adj < .05) %>%
    dplyr::mutate(type = paste0(ident.1,".vs.",ident.2), trans = avg_logFC > 0) %>%
    dplyr::arrange(desc(avg_logFC))
  x$trans = factor(x$trans, labels = c("down", "up"))
  return(x)
}, 
ident.1 = rep(paste0("aDNT_rep",1:2),2),
ident.2 = rep(paste0("nDNT_rep",1:2),2),
subset.ident = c(0,0,1,1), 
SIMPLIFY = F)
head(df[[1]])
trans.marker = mapply(function(df){
  x = dplyr::inner_join(df[[1]], df[[2]], by = c("gene", "trans"), suffix = c(".rep1", ".rep2"))
}, SIMPLIFY = F, df = list("DNT0" = df[1:2], "DNT1" = df[3:4]))
head(trans.marker[[1]])
save(trans.marker, file = "01.trans.marker.Rdata")
```
# cca unfilter
```{r}
plan("sequential")
names(sct.dnt)
cca.unfilter = mapply(function(object.list){
  anchor.features = SelectIntegrationFeatures(object.list = object.list, 
                                              nfeatures = 3000) 
  object.integrated = PrepSCTIntegration(object.list = object.list, 
                                         anchor.features = anchor.features) %>%
    lapply(FUN = RunPCA, features = anchor.features, verbose = F) %>%
    FindIntegrationAnchors(object.list = ., 
                           reduction = "rpca",
                           anchor.features = anchor.features, 
                           normalization.method = "SCT",
                           reference = 1) %>% # long time
    IntegrateData(anchorset = ., 
                  normalization.method = "SCT") %>%
    RunPCA(object = .) %>%
    RunUMAP(object = ., 
            dims = 1:30) %>%
    FindNeighbors(dims = 1:30) %>%
    FindClusters(resolution = .05)
  ##### scale for finding markers #####
  DefaultAssay(object.integrated) = "RNA"
  object.integrated = NormalizeData(object.integrated,
                                    normalization.method = "LogNormalize",
                                    scale.factor = 10000)
  object.integrated = FindVariableFeatures(object.integrated,
                                           selection.method = "vst",
                                           nfeatures = 2000)
  object.integrated = ScaleData(object.integrated,
                                vars.to.regress = "percent.mt")
  return(object.integrated)
}, 
object.list = list("nDNT" = sct.dnt[5:6], "aDNT" = sct.dnt[1:2], "DNT" = sct.dnt[c(5:6, 1:2)]), SIMPLIFY = F)
# object = cca.unfilter$DNT
# DefaultAssay(object) = "integrated"
# object = FindClusters(object, resolution = .02)
# DimPlot(object)
save(cca.unfilter, file = "01.cca.unfilter.Rdata")
```
```{r modify the metadata}
meta.all = mapply(function(object){
  ##### gene expression #####
  DefaultAssay(object) = "RNA"
  meta.gene = FetchData(object = object, 
                        vars = c("Cd3d", "Cd3e", "Cd3g", "Cd4", "Cd8a", "Cd8b1", "Klrb1c"), 
                        slot = "data") %>%
    tibble::rownames_to_column("cell.rename")
  head(meta.gene)
  ##### save the pre metadata #####
  metapre = object@meta.data
  ##### merge pre and modified metadata #####
  meta.all = metapre %>%
    tibble::rownames_to_column("cell.rename") %>%
    dplyr::left_join(x = ., y = meta.gene, by = "cell.rename") %>%
    tibble::column_to_rownames("cell.rename")
  head(meta.all)
  ##### T identity #####
  if (("Cd4" %in% colnames(meta.all)) == F){
    meta.all$DNTid = (meta.all$Cd3d > 0 |meta.all$Cd3e > 0| meta.all$Cd3g > 0) &
      meta.all$Klrb1c == 0 & meta.all$Cd8b1 == 0
  } else {
    meta.all$DNTid = (meta.all$Cd3d > 0 |meta.all$Cd3e > 0| meta.all$Cd3g > 0) &
      meta.all$Cd4 == 0 & meta.all$Klrb1c == 0 & meta.all$Cd8b1 == 0
  }
  
  table(meta.all$DNTid, meta.all$integrated_snn_res.0.05)
  meta.all$contaminate = meta.all$DNTid == FALSE
  table(meta.all$contaminate, meta.all$integrated_snn_res.0.05)
  meta.all$contaminate = factor(meta.all$contaminate, labels = c("kept", "removed"))
  return(meta.all)
}, object = cca.unfilter, SIMPLIFY = F)
save(meta.all, file = "01.cca.unfilter.meta.all.Rdata")
```
## markers
```{r}
cc.marker = mapply(function(object){
  print(table(object$seurat_clusters, object$orig.ident))
  lapply(sort(unique(object$seurat_clusters)), function(i){
    if (min(table(object$orig.ident[object$seurat_clusters == i])) > 30){
      FindConservedMarkers(object, ident.1 = i, grouping.var = "orig.ident", 
                           assay = "RNA", slot = "data", only.pos = T) %>%
        tibble::rownames_to_column("gene") %>%
        dplyr::filter(max_pval < .05 & minimump_p_val < .05) %>%
        dplyr::mutate(cluster = i)
    }}) %>% 
    data.table::rbindlist() %>%
    dplyr::left_join(y = FindAllMarkers(object, assay = "RNA", slot = "data", only.pos = T), by = c("cluster", "gene")) %>%
    dplyr::group_by(cluster) %>%
    dplyr::arrange(desc(avg_logFC), .by_group = T)
}, object = cca.unfilter, SIMPLIFY = F)
names(cc.marker)
save(cc.marker, file = "01.cca.unfilter.cc.marker.Rdata")
```


# DoubletFinder and DoubletDecon
## DoubletDecon use sct results
```{r}
source("../26.MAIT/DoubletDecon.f.r")
############################################################
# remove doublets using DoubletDecon, dataset need to be clustered first, must have more than two clusters, the plot window should be the maximized.
############################################################
dev.off()
object = lapply(sct.kept[5:6], function(i){
  i = i %>% 
    RunPCA() %>% 
    RunUMAP(dims = 1:30) %>% 
    FindNeighbors() %>% 
    FindClusters(resolution = .1)})
nDNT.decon = mapply(DoubletDecon.f, 
                    object = object, 
                    filename = c("nDNT_rep1.sct", "nDNT_rep2.sct"), 
                    rhop = 1,
                    SIMPLIFY = F)
save(nDNT.decon, file = "01.nDNT.decon.Rdata")
```
## DoubletFinder union Decon use sct results
```{r}
source("../26.MAIT/DoubletFinder.f.r")
doublet.finder.res = mapply(DoubletFinder.f, object = nDNT.decon, SIMPLIFY = F)
save(doublet.finder.res, file = "01.DoubletDecon.DoubletFinder.Rdata")
```

